Fast Radio Burst Cosmology

### 2014-04-10 writeup/dm_angular_ps.tex
Work on the estimator.

%\documentclass[onecolumn,amsmath,amssymb,floatfix,12pt,nofootinbib]{revtex4}
\documentclass[twocolumn,prd,noshowpacs,nofootinbib,amsmath,amssymb]{revtex4}
\usepackage{graphicx,epsfig,psfrag,bm,amssymb}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{mciteplus}
\usepackage{color}
\usepackage{mathrsfs,amsfonts,hepunits, color}
\usepackage{mciteplus}
\usepackage{tikz}
\usepackage{slashed}
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{trees}
\usetikzlibrary{matrix,arrows} 				% For commutative diagram
											% http://www.felixl.de/commu.pdf
\usetikzlibrary{positioning}				% For "above of=" commands
\usetikzlibrary{calc,through}				% For coordinates
\usetikzlibrary{decorations.pathreplacing}  % For curly braces
\usepackage{pgffor}							% For repeating patterns

@@

@@

@@

For some line of sight originating from redshift $z$

\begin{equation}
{\rm DM}^{z}(\hat{n}) = \int_0^{\eta^z} d\eta \frac{n_e(\eta\hat{n})}{(1+z(\eta))^2}
\end{equation}
is the dispersion-measure random field in angular direction $\hat{n}$ where
\begin{equation}
n_e(\eta^z\hat{n}) \equiv n_e(z,\hat{n}) =  \bar{n}_{e0} (1+z)^3 \left[1+\delta_e(z,\hat{n})\right]
\end{equation}
is the electron distribution of the Universe and $\bar{n}_{e0}$ is the mean cosmological density of free electrons today, and $\delta_e(z,\hat{n})$ is 
the fractional overdensity of electrons.
We define the dispersion-measure distance (or dispersion distance), the equivalent distance required to
produce a given amount of dispersion in an observed radio signal in the Universe today, as 
$ D = {{\rm DM}}/{\bar{n}_{e0}}$

>>>>

%\documentclass[onecolumn,amsmath,amssymb,floatfix,12pt,nofootinbib]{revtex4}
\documentclass[twocolumn,prd,noshowpacs,nofootinbib,amsmath,amssymb]{revtex4}
\usepackage{graphicx,epsfig,psfrag,bm,amssymb}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{mciteplus}
\usepackage{color}
\usepackage{mathrsfs,amsfonts,hepunits, color}
\usepackage{mciteplus}
\usepackage{tikz}
\usepackage{slashed}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{trees}
\usetikzlibrary{matrix,arrows} 				% For commutative diagram
											% http://www.felixl.de/commu.pdf
\usetikzlibrary{positioning}				% For "above of=" commands
\usetikzlibrary{calc,through}				% For coordinates
\usetikzlibrary{decorations.pathreplacing}  % For curly braces
\usepackage{pgffor}							% For repeating patterns

@@

\author{Kiyoshi Wesley Masui}
 \affiliation{Department of Physics and Astronomy, University of British 
Columbia, Vancouver, BC, Canada V6T 1Z1}

@@

\subsection{Dispersion measure and PDF}

@@

The dispersion measure of a FRB observed in some angular direction $\hat{n}$ and
originating from redshift $z$ is
\begin{equation}
{\rm DM}^{z}(\hat{n}) = \int_0^{\eta^z} d\eta
\frac{n_e(\eta\hat{n})}{(1+z(\eta))^2}.
\end{equation}
Here, $\eta$ is the conformal distance as a function of redshift, and
$n_e(\eta^z\hat{n})$ is the three dimensional electron density.  We choose to
separate $n_e$ into four components as follows:
\begin{equation}
\label{e:ne_sep}
n_e(\hat{n}, z) = n_e^{\rm gal}(\hat{n}, z) + \bar{n}_e(z) +
    \bar{n}_e(z)\delta_e(\hat{n}, z) + n_e^{\rm src}(\hat{n}, z).
\end{equation}
Here $n_e^{\rm gal}(\hat{n}, z)$ is the contribution to the electron density
from the Milky Way galaxy and other local structures and only contributes at
low redshift. $\bar{n}_e(z)$ is the mean electron density is and related to the
current day electron density by $\bar{n}_e = \bar{n}_{e0} (1+z)^3$. 
$\delta_e(z,\hat{n})$ is the large-scale fractional over-density of electrons.
$n_e^{\rm src}(\hat{n}, z)$ are the electrons associated with the FRB source,
which only contributes as $z \approx z^{\rm src}$ and accounts for two effects.
First the FRB itself
may have some intrinsic dispersion measure due to the environment of the
progenitor.  Second, it is unlikely that FRB events are uniformly distributed
throughout the Universe and they presumably trace matter in some way. When
considering statics of FRB dispersion measures it is necessary to account for
this bias.

We define the dispersion-measure distance (or dispersion distance),
the equivalent distance required to
produce a given amount of dispersion in an observed radio signal in the
Universe today, as $ D = {{\rm DM}}/{\bar{n}_{e0}}$
We separate the dispersion distance into the contributions from each
term in \ref{e:ne_sep}:
\begin{equation}
\label{e:D_sep}
D = D^{\rm gal} + D^{\bar{n}_e} + D^{\delta_e}
         + D^{\rm src}.
\end{equation}
For a population of FRBs originating at cosmological distances, ${\rm DM}^{\rm
src}$ is only a single function of $\hat{n}$ which can presumably be well
measured.  In addition, this function could in principal be well characterized
by independent data such as observations of pulsars.  As such, we will
here-after we ignore this term.

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\{D_i\}$ with known angular positions but with unknown
redshifts depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\{z_i\})$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\{z_i\}) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $\{D^{\bar{n_e}}_i\}$ are
uncorrelated:
\begin{equation}
\langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
    - \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle
    = \delta_{ij} \times \textrm{(something to do with $p(z_i)$)}
\end{equation}

We further assume that $\{D^{\rm src}_i\}$ are uncorrelated. Again this may not
be strictly true due to clustering of FRB sources and again this may bias our
estimator. We parameterized the variance from this term as 
\begin{equation}
    \langle D^{\rm src}_i D^{\rm src}_j \rangle
    - \langle D^{\rm src}_i \rangle \langle D^{\rm src}_j \rangle
    = \delta_{ij} \sigma^2_{\rm src}
\end{equation}

We further assume that the terms in Equation \ref{e:D_sep} are mutually uncorrelated

\begin{comment}
is the dispersion-measure random field in angular direction $\hat{n}$ where
\begin{equation}
n_e(\eta^z\hat{n}) \equiv n_e(z,\hat{n}) =  \bar{n}_{e0} (1+z)^3 \left[1+\delta_e(z,\hat{n})\right]
\end{equation}
is the electron distribution of the Universe and $\bar{n}_{e0}$ is the mean cosmological density of free electrons today, and $\delta_e(z,\hat{n})$ is 
the fractional over-density of electrons.
\end{comment}

\subsection{Relating dispersion measure angular power-spectrum to cosmology}
